### 2024-02-12 outputs/paper/paper.qmd
include simulation

---
title: "Misperceived Social Norms about Women Working Outside the Home and Its Effect on Women in Saudi Arabia"
author: 
  - Boxuan Yi
  - Ruoxian Wu
thanks: "Code and data are available at: https://github.com/ScarletWu/Misperceived-Social-Norms-Women-Working-Outside-the-Home-in-Saudi-Arabia. Replication on Social Science Reproduction platform is available at: https://www.socialsciencereproduction.org/reproductions/1557/"
date: "February 12, 2024"
date-format: long
abstract: "Husbands who hold guardianship and discourage their wives from working outside the home are often perceived as the main reason for the low female workforce participation in Saudi Arabia. Bursztyn, González, and Yanagizawa-Drott explore this social norm through surveys, experiments and visualizations, revealing that a majority of young husbands in Saudi Arabia privately support women working outside the home but underestimate the support of other men, and correcting this misperception increases husbands’ willingness to help wives search for jobs. This paper replicates the data from their study to analyze young husbands’ actual opinions and perceptions, confirming similar findings. This paper also discusses strategies for reducing misperceptions and the unusual Saudi female employment rate during the pandemic."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| message: false
#| echo: false
# Install packages we need
install.packages('haven')
install.packages('dplyr')
install.packages('ggplot2')
install.packages('readr')
install.packages('here')
install.packages('janitor')
install.packages('knitr')
install.packages('kableExtra')
install.packages('tidyverse')
```

```{r}
#| message: false
#| echo: false
library(haven)
library(dplyr)
library(ggplot2)
library(readr)
library(here)
library(janitor)
library(knitr)
library(kableExtra)
library(tidyverse)
```

# Introduction
Barriers to the employment of women exert a significant influence on female labor, presenting a constraint to increasing economic opportunities for women [@sciencedirectOrganizationalEconomic]. Saudi Arabia is a recognizable example of a country with a low rate of female labor force participation. In 2022, the female labor force participation rate, defined as the percentage of the female population aged 15 and above that is employed, is only 27.8% in Saudi Arabia. This is notably lower than the global average of 47.3%, while the male labor force participation rate in Saudi Arabia remains high at 80% [@worldbankLaborForce].

The prevailing societal norms in Saudi Arabia, characterized by a low rate of women working outside the home and the customary practice of husbands' guardianship with decision-making authority, contribute to the perception that Saudi men are unwilling to permit their wives to work [@MisperceivedSocialNorm]. The paper ‘Misperceived Social Norms: Women Working Outside the Home in Saudi Arabia’ by Leonardo Bursztyn, Alessandra L. González, and David Yanagizawa-Drott delves into this social norm, exploring Saudi men’s perceptions of opinions regarding WWOH.

Based on survey responses and experiments results, they find that a large portion of young married men in Saudi Arabia privately support women working outside the home. Importantly, these men greatly underestimate the proportion of support among other similar men [@MisperceivedSocialNorm]. The study population is young married husbands aged 18 to 35 in Saudi Arabia. We aim to replicate the estimand that focuses on the effects of providing husbands with the true proportion of support. Their findings suggest that men who are given the actual proportion of their peers supporting WWOH are more likely to support their wives’ labor supply, and their wives are more likely to apply for job, attend interviews, be employed and enrol in driving lessons. Correcting misperceived beliefs yields positive impacts on females employment, and insufficient communication is associated with an increased likelihood of misperceptions [@MisperceivedSocialNorm].

Our paper will follow a reproduction of Leonardo Bursztyn, Alessandra L. González, and David Yanagizawa-Drott’s findings. We seek to replicate two claims, (1) There is a wedge between the actual and perceived proportion of men supporting women working outside the home, and (2) Most Saudi men who privately support WWOH fail to understand that others do as well, and they are more likely to help their wives find jobs when they know the actual portion of men supporting WWOH. The experiments and surveys results we will try to replicate are (1) the actual portion of support of WWOH from young husbands, (2) the existence of misperception, and (3) the effects of correcting beliefs on female employment in Saudi Arabia. Our reproduction used the programming language R [@r], the analysis used the following packages: Haven[@rHaven], Dplyr [@rDplyr], Ggplot2 [@rGgplot2], Readr [@rReadr], Here [@rHere], Janitor [@rJanitor], KableExtra [@rKableExtra], Knitr [@rKnitr], Tidyverse [@rTidyverse].

In the next section, we will discuss data including data source, methodology and features. After that, we will create a reproduction of the selected results to verify their data and their findings based on the data. In the last section, we will discuss the impact of informational intervention, the use of social proof to overcome harmful norms, and the inverse pattern of female labor force participation rate in Saudi Arabia during the pandemic. We will also conclude ethics, potential bias, and weaknesses topics, as well as a push for future research.

# Data
## Source
The paper used for replication is from American Economic Association, American Economic Review. The replication package is downloaded from Misperceived Social Norms: Women Working Outside the Home in Saudi Arabia webpage, under Additional Materials.

Our reproduction seeks to address two findings written in the original paper, using results they collected from the main experiment, its follow-up and online survey.

## Methodology and Features
### Main experiment and Follow-up
The original study recruited 500 married Saudi males aged 18 to 35 from Riyadh as participants. Initial contacts were established using a recruiter database, with further recruitment using a combination of snowball sampling and random street intercept [@MisperceivedSocialNorm].

Participants were organized into 17 sessions, each comprising 30 individuals. The survey, administered through Qualtrics, encompassed questions about demographic information, opinions, agreement with statements, and estimations regarding the agreement of other  29 participants with each statement. Particularly, there were two statements about female labor participation. Following this, half of the participants were assigned to a treatment group randomly. Those in the treatment group were provided with information on the proportion of respondents who agreed or disagreed with each statement, while participants in the control group received no information. Respondents were then asked to make an incentivized choice between a five-dollar Amazon gift card and the opportunity to sign up their wives for a job matching platform [@MisperceivedSocialNorm]. Follow-up calls were conducted 3-5 months after the main experiment. These calls collected information about whether the wife was employed during the main experiment, whether the wife was currently employed, and if she was currently employed, whether the job was outside the home [@MisperceivedSocialNorm].

The main experiment dataset contains 500 observations and 36 variables. Variables include age, level of education, number of children, whether this man and his wife are employed now, number of people known in the group and other variables related to WWOH statements. The follow-up dataset contains similar demographic data and information about wives’ employment, interviews and driving lessons. There are 33 variables in the follow-up dataset. Details and explanations using visualization will be further discusses in the next section.

### Online survey
The online survey included approximately 1500 married Saudi men aged 18 to 35, randomly assigned to either a treatment or control group. In the control group, participants were presented with statements that were contentious but not stigmatized. In the treatment group, participants received an identical list of items with an additional, potentially stigmatized item for which the experimenters would like to elicit beliefs. All participants were then asked to indicate the number of statements they agree with, and the respondents from the control group were asked if they agreed with a WWOH statement individually [@MisperceivedSocialNorm]. Finally, the experimenters asked participants in the control condition to estimate the percentage of other participants who reported agreeing with the same statement, and asked participants from the treatment group to estimate the percentage of other participants who would privately agree with the WWOH statement [@MisperceivedSocialNorm].

The dataset resulting from online survey comprises 1460 observations and 21 variables. These variables include condition, age, employed_now, employed_wife, employed_out_wife, children, and the responses to the survey questions for both groups as indicated above, along with participants’ confidence levels in some of the answers. Details and explanations using visualization will be further discusses in the next section.

# Results
```{r}
#| message: false
#| echo: false
# Read the dataset
main_clean <- read_dta(here("inputs/data/01_main_exp_clean_original.dta"))
follow_up_clean <- read_dta(here("inputs/data/02_follow_up_clean_original.dta"))
online_survey_clean <- read_dta(here("inputs/data/03_1st_online_svy_clean_original.dta"))
```

```{r}
#| include: false
# Define the path
datadir <- "/cloud/project/inputs/data/"
data_path <- file.path(datadir, "clean_data/01_main_exp_clean.dta")

# Load the dataset
df <- read_dta(data_path)

# Define the variable prefix
q <- "outside"

# Create new variables
df[[paste0(q, "_guess_share")]] <- df[[paste0(q, "_guess")]] / 30
df[[paste0(q, "_others_truth")]] <- ifelse(df$stag == 1, df[[paste0(q, "_objective")]] / 30, NA)
q_self_mean <- mean(df[[paste0(q, "_self")]], na.rm = TRUE)
df[[paste0(q, "_wedge_share")]] <- (df[[paste0(q, "_wedge")]] / 30) * 100

```

```{r}
#| label: fig-1
#| fig-cap: Histogram of Perceptual Gaps in Beliefs About Women Working Outside the Home
#| echo: false
#| warning: false
#| message: false

# Plotting a histogram of the wedge share
f2 <- ggplot(df, aes(x = .data[[paste0(q, "_wedge_share")]])) +
  geom_histogram(binwidth = 10.5, fill = "lightgrey", color = "black") + 
  scale_x_continuous("Wedge (guess % - objective %)", 
                     breaks = seq(-100, 100, by = 20), 
                     limits = c(-100, 100), 
                     expand = c(0.1, 0.5)) +
  scale_y_continuous("Density", expand = c(0.5, 0)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "#FF4136", size = 0.5) + # Less prominent zero line
  labs(title = "Distribution of Wedge (Guess % - Objective %)", # Add a title
       caption = "Red dashed line represents the zero point") + # Add a caption
  theme_minimal(base_size = 12) + 
  theme(
    plot.background = element_blank(),
    panel.background = element_rect(fill = "white", colour = "black"), 
    axis.text = element_text(size = 12, color = "black"), 
    axis.title = element_text(size = 14, color = "black"),
    panel.grid.major = element_line(color = "lightgrey", size = 0.2), 
    panel.grid.minor = element_blank(),
    plot.title = element_text(hjust = 0.5), 
    plot.caption = element_text(hjust = 0.5) 
  )
f2
```

The authors calculated "wedges" which represent the difference between what participants guessed and the actual percentage of agreement among their session peers regarding women working outside the home (WWOH). The negative wedge indicates that the participant underestimated the support for WWOH. Conversely, a positive wedge suggests that the participant overestimated the support for WWOH. We replicate the histograms to check the distribution of the wedges between participants' beliefs about the support for WWOH among their session participants and the actual levels of support. @fig-1 specifically maps the distribution of these "wedges" -- the differences between each participant's guess and the actual percentage of session participants who agreed with the pro-WWOH statement. It was found that more than half of participants underestimated the support for WWOH. There is a clear gap between perceived and actual social norms regarding women's work outside the home. The left skew of the histogram would indicate that a majority of participants underestimated the true level of support for WWOH.

By using the regression analysis, we examine the treatment effect on the participants' likelihood to sign up for a job matching service for their wives. The treatment variable is significant across all model specifications, with coefficients ranging from 0.0853 to 0.0899. The constant term varies considerably across different model specifications, suggesting differing baseline propensities for signing up absent the treatment. The p-values for the treatment effect, under various robustness checks such as robust standard errors, wild bootstrap, and permutation test, all indicate significance, ranging from 0.008 to 0.038. These results reinforcing the treatment's positive impact on sign-up rates (@tbl-summary_table).

```{r}
#| include: false
# Observations
total_rows <- nrow(main_clean)
c_rows <- sum(main_clean$condition2 == 0, na.rm = TRUE)
t_rows <- sum(main_clean$condition2 == 1, na.rm = TRUE)

# Age
total_age <- round(mean(main_clean$age, na.rm = TRUE), 2)
c_age <- round(mean(main_clean$age[main_clean$condition2 == 0], na.rm = TRUE), 2)
t_age <- round(mean(main_clean$age[main_clean$condition2 == 1], na.rm = TRUE), 2)
# Age standard deviation
total_age_sd <- paste0("(", round(sd(main_clean$age, na.rm = TRUE), 2), ")")
c_age_sd <- paste0("(", round(sd(main_clean$age[main_clean$condition2 == 0], na.rm = TRUE), 2), ")")
t_age_sd <- paste0("(", round(sd(main_clean$age[main_clean$condition2 == 1], na.rm = TRUE), 2), ")")

# Number of children
total_ch <- round(mean(main_clean$children, na.rm = TRUE), 2)
c_ch <- round(mean(main_clean$children[main_clean$condition2 == 0], na.rm = TRUE), 2)
t_ch <- round(mean(main_clean$children[main_clean$condition2 == 1], na.rm = TRUE), 2)
# Number of children standard deviation
total_ch_sd <- paste0("(", round(sd(main_clean$children, na.rm = TRUE), 2), ")")
c_ch_sd <- paste0("(", round(sd(main_clean$children[main_clean$condition2 == 0], na.rm = TRUE), 2), ")")
t_ch_sd <- paste0("(", round(sd(main_clean$children[main_clean$condition2 == 1], na.rm = TRUE), 2), ")")

# College Degree (%)
total_cd <- round(mean(main_clean$college_deg == 1, na.rm = TRUE) * 100, 2)
c_cd <- main_clean |> filter(condition2 == 0) |>
  summarise(c_cd = round(mean(college_deg == 1, na.rm = TRUE) * 100, 2)) |> pull(c_cd)
t_cd <- main_clean |> filter(condition2 == 1) |>
  summarise(t_cd = round(mean(college_deg == 1, na.rm = TRUE) * 100, 2)) |> pull(t_cd)

# Employed (%)
total_emp <- round((sum(main_clean$employed_now == 1, na.rm = TRUE) / total_rows) * 100, 2)
c_emp <- main_clean %>%
  filter(condition2 == 0) %>%
  summarise(c_emp = round(sum(employed_now == 1, na.rm = TRUE) / c_rows * 100, 2)) %>%
  pull(c_emp)
t_emp <- main_clean %>%
  filter(condition2 == 1) %>%
  summarise(t_emp = round(sum(employed_now == 1, na.rm = TRUE) / t_rows * 100, 2)) %>%
  pull(t_emp)

# Wife employed (%)
total_we <- round((sum(main_clean$employed_wife == 1, na.rm = TRUE) / total_rows) * 100, 2)
c_we <- main_clean %>%
  filter(condition2 == 0) %>%
  summarise(c_we = round(sum(employed_wife == 1, na.rm = TRUE) / c_rows * 100, 2)) %>%
  pull(c_we)
t_we <- main_clean %>%
  filter(condition2 == 1) %>%
  summarise(t_we = round(sum(employed_wife == 1, na.rm = TRUE) / t_rows * 100, 2)) %>%
  pull(t_we)

# Wife working outside the home (% retrospective follow-up)
c_wo <- follow_up_clean %>%
  filter(condition2 == 0) %>%
  filter(matched == 1) %>%
  summarise(c_wo = round(mean(employed_3mos_out_fl) * 100, 2)) %>%
  pull(c_wo)
t_wo <- follow_up_clean %>%
  filter(condition2 == 1) %>%
  filter(matched == 1) %>%
  summarise(t_wo = round(mean(employed_3mos_out_fl) * 100, 2)) %>%
  pull(t_wo)
total_wo <- follow_up_clean %>%
  filter(matched == 1) %>%
  summarise(total_wo = round(mean(employed_3mos_out_fl) * 100, 2)) %>%
  pull(total_wo)

# Other participants known (%)
total_kn <- round(mean(main_clean$num_know_per, na.rm = TRUE) * 100, 2)
c_kn <- main_clean %>%
  filter(condition2 == 0) %>%
  summarise(c_kn = round(mean(num_know_per, na.rm = TRUE) * 100, 2)) %>%
  pull(c_kn)
t_kn <- main_clean %>%
  filter(condition2 == 1) %>%
  summarise(t_kn = round(mean(num_know_per, na.rm = TRUE) * 100, 2)) %>%
  pull(t_kn)

# Other participants known standard deviation (%)
total_kn_sd <- paste0("(", round(sd(main_clean$num_know_per, na.rm = TRUE) * 100, 2), ")")
c_kn_sd <- paste0("(", round(sd(main_clean$num_know_per[main_clean$condition2 == 0], na.rm = TRUE) * 100, 2), ")")
t_kn_sd <- paste0("(", round(sd(main_clean$num_know_per[main_clean$condition2 == 1], na.rm = TRUE) * 100, 2), ")")

# Other participants with mutual friends (%)
total_mf <- round(mean(main_clean$num_mfs_per, na.rm = TRUE) * 100, 2)
c_mf <- main_clean %>%
  filter(condition2 == 0) %>%
  summarise(c_mf = round(mean(num_mfs_per, na.rm = TRUE) * 100, 2)) %>%
  pull(c_mf)
t_mf <- main_clean %>%
  filter(condition2 == 1) %>%
  summarise(t_mf = round(mean(num_mfs_per, na.rm = TRUE) * 100, 2)) %>%
  pull(t_mf)

# Other participants with mutual friends standard deviation (%)
total_mf_sd <- paste0("(", round(sd(main_clean$num_mfs_per, na.rm = TRUE) * 100, 2), ")")
c_mf_sd <- paste0("(", round(sd(main_clean$num_mfs_per[main_clean$condition2 == 0], na.rm = TRUE) * 100, 2), ")")
t_mf_sd <- paste0("(", round(sd(main_clean$num_mfs_per[main_clean$condition2 == 1], na.rm = TRUE) * 100, 2), ")")
```

```{r}
#| include: false
# Assign groups
total_group <- c(total_rows, total_age, total_age_sd, total_ch, total_ch_sd, total_cd, total_emp, total_we, total_wo, total_kn, total_kn_sd, total_mf, total_mf_sd)
control_group <- c(c_rows, c_age, c_age_sd, c_ch, c_ch_sd, c_cd, c_emp, c_we, c_wo, c_kn,c_kn_sd, c_mf, c_mf_sd)
treatment_group <- c(t_rows, t_age, t_age_sd, t_ch, t_ch_sd, t_cd, t_emp, t_we, t_wo, t_kn, t_kn_sd, t_mf, t_mf_sd)
```

```{r}
#| message: false
#| echo: false
#| label: tbl-summary_table
#| tbl-cap: Summary Statistics (Main Experiment)
table1 <- cbind(total_group, control_group, treatment_group)
row.names(table1)= c("Observations","Age","","Number of Children","","College Degree (%)","Employed (%)","Wife Employed (%)", "Wife Working Outside the Home (% retrospective follow-up)", "Other Participants Known (%)", "","Other Participants with Mutual Friends (%)","")

table1 |> kbl(booktabs = T, col.names = c("All","Control","Treatment"), align=rep('c', 4)) |> 
  row_spec(1, hline_after = TRUE) |> kable_styling()
```

```{r}
#| message: false
#| echo: false
p_value_figure3 <- chisq.test(table(main_clean$condition2, main_clean$signed_up_number))

percentage_data_figure3 <- main_clean |>
  mutate(condition2 = ifelse(condition2 == 0, 'Control', 'Treatment')) |>
  group_by(condition2) |>
  summarise(percentage = mean(signed_up_number == 1) * 100)
```

```{r}
#| message: false
#| echo: false
#| label: fig-job_sign_up
#| fig-cap: Job Matching Service Sign Up (Main Experiment)
plot_figure3 <- ggplot(percentage_data_figure3, aes(x = condition2, y = percentage)) +
  geom_bar(stat = "identity", fill = "grey", color = "black", alpha = 0.9) +
  labs(x = "",
       y = "Percent sign up") +
  geom_text(aes(label = paste0(round(percentage, 2), "%")),
            position = position_dodge(width = 0.9),
            vjust = -0.5) +
  annotate("text", x = 1.5, y = 37, label = sprintf("p-value = %.4f", p_value_figure3$p.value)) +
  theme_minimal() + 
  ylim(0, 40)

plot_figure3
```

@fig-job_sign_up illustrates the difference in job matching service sign-up rates between the control and treatment groups. The bar for the treatment group would be higher than the control group's, showing a clear increase in the likelihood of signing up due to the treatment. The specific values of 23.48% for the control group and 32.02% for the treatment group, along with a p-value of 0.0424, would be represented, indicating a statistically significant effect of the information intervention on the participants' actions.

The combination of @tbl-summary_table and @fig-job_sign_up demonstrates the efficacy of the information treatment in altering the participants' behavior towards women's labor market participation, revealing not only a statistical significance but also practical importance in the context of social norms and labor supply outcomes in Saudi Arabia.

\newpage
The the longer-term outcomes of the treatment are shown in @fig-3 with error bars indicating the 95% confidence intervals for these proportions.

The job matching service sign-up rates is presented in @fig-3 a. In the control group, 6.04% of participants chose to sign their wives up for the job service instead of taking a gift card payment while 16.74% participants in the treatment group chose the job service (p-value = 0.001). The increased sign-up rate in the treatment group more young men accepted the idea of WWOH after the correction of misperceptions.

In @fig-3 b, the informational intervention has impact on participants' wives were interviewed for a job outside the home. After the intervention, there was a statistically significant increase in interviews, with the percentage of wives who interviewed for a job outside the home rising from 1.1% to 5.95%. This more than fivefold increase (p-value = 0.013) suggests that the treatment had a notable effect on the interview rates for jobs outside the home.

Similarly, the informational intervention effects on the employment rate of participants' wives in jobs outside the home according to @fig-3 c. The employment rate post-intervention increased from 7.69% to 9.73%. Although this represents an increase, it is not statistically significant (p-value = 0.456), implying that while the intervention might have had an encouraging effect on job-seeking behavior, it did not translate into a significant difference in actual employment within the study period.

Besides, @fig-3 d indicates that the change in husbands' willingness to sign their wives up for driving lessons as a result of the informational intervention. There is a significant increase from 68.13% to 78.38% in the proportion of husbands who reported they would sign their wives up for driving lessons (p-value = 0.03). This result indicates that the informational intervention not only had an impact on labor supply outcomes but also potentially affected broader social norms and attitudes toward women's rights and autonomy in Saudi Arabia, as evidenced by the increased openness to women's mobility through driving.

Overall, the treatment has a significant impact on both immediate and longer-term outcomes related to women's labor market participation and related behaviors (@fig-3).

```{r}
#| include: false
# Define path 
datadir <- "/cloud/project/inputs/data/"
data_path <- file.path(datadir, "clean_data/02_follow_up_clean.dta")

# Load the dataset
df <- read_dta(data_path)

# Keep matched responses and rename variable if necessary
df_filtered <- df %>%
  filter(matched == 1)

# Calculate rates and their confidence intervals
df_stats <- df_filtered %>%
  mutate(condition = if_else(is.na(condition), 0, condition)) %>%
  group_by(condition) %>%
  summarise(
    n = n(),
    applied_rate = mean(applied_out_fl, na.rm = TRUE) * 100,
    interviewed_rate = mean(interviewed_out_fl, na.rm = TRUE) * 100,
    employed_now_rate = mean(employed_now_out_fl, na.rm = TRUE) * 100,
    driving_rate = mean(driving_fl, na.rm = TRUE) * 100 
  ) %>%
  mutate(
    applied_se = sqrt((applied_rate / 100) * (1 - (applied_rate / 100)) / n) * 100,
    applied_rate_lo = applied_rate - 1.96 * applied_se,
    applied_rate_hi = applied_rate + 1.96 * applied_se,
    interviewed_se = sqrt((interviewed_rate / 100) * (1 - (interviewed_rate / 100)) / n) * 100,
    interviewed_rate_lo = interviewed_rate - 1.96 * interviewed_se,
    interviewed_rate_hi = interviewed_rate + 1.96 * interviewed_se,
    employed_now_se = sqrt((employed_now_rate / 100) * (1 - (employed_now_rate / 100)) / n) * 100,
    employed_now_rate_lo = employed_now_rate - 1.96 * employed_now_se,
    employed_now_rate_hi = employed_now_rate + 1.96 * employed_now_se,
    driving_se = sqrt((driving_rate / 100) * (1 - (driving_rate / 100)) / n) * 100, 
    driving_rate_lo = driving_rate - 1.96 * driving_se,  
    driving_rate_hi = driving_rate + 1.96 * driving_se   
  ) %>%
  select(-n)

# p values
p_value_applied <- t.test(applied_out_fl ~ factor(condition), data = df_filtered, na.rm = TRUE)$p.value
p_value_interviewed <- t.test(interviewed_out_fl ~ factor(condition), data = df_filtered, na.rm = TRUE)$p.value
p_value_employed_now <- t.test(employed_now_out_fl ~ factor(condition), data = df_filtered, na.rm = TRUE)$p.value
p_value_driving <- t.test(driving_fl ~ factor(condition), data = df_filtered, na.rm = TRUE)$p.value

```

```{r}
#| label: fig-3
#| fig-cap: Histogram of Perceptual Gaps in Beliefs About Women Working Outside the Home
#| echo: false
#| warning: false
#| message: false
#| layout-ncol: 2
#| fig-subcap: ["Applied Rate","Interviewed Rate", "Employed Now Rate", "Driving Rate"]

# Applied Rate
applied_rate_4a <- ggplot(df_stats, aes(x = factor(condition), y = applied_rate)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black", width = 0.7) +
  geom_errorbar(aes(ymin = applied_rate_lo, ymax = applied_rate_hi), width = 0.2, color = "black") +
  geom_text(aes(label = sprintf("%.2f%%", applied_rate), y = applied_rate + 1), position = position_dodge(width = 0.7), vjust = 0) +
  annotate("text", x = 1, y = max(df_stats$applied_rate_hi) + 3, label = paste("p-value =", round(p_value_applied, 3)), size = 4) +
  labs(x = "", y = "% Applied", title = "Long-Term Labor Supply Outcomes") +
  theme_minimal() +
  scale_x_discrete(labels = c("Control", "Treatment"))

applied_rate_4a

# Interviewed Rate

interviewed_rate_4b <- ggplot(df_stats, aes(x = factor(condition), y = interviewed_rate)) +
  geom_bar(stat = "identity", fill = "yellow", color = "black", width = 0.7) +
  geom_errorbar(aes(ymin = interviewed_rate_lo, ymax = interviewed_rate_hi), width = 0.2, color = "black") +
  geom_text(aes(label = sprintf("%.2f%%", interviewed_rate), y = interviewed_rate_hi + 2), vjust = 0) +
  annotate("text", x = 1.5, y = max(df_stats$interviewed_rate_hi) + 5, label = paste("p-value =", round(p_value_interviewed, 3)), size = 4) +
  labs(y = "% Interviewed", x = "") +
  theme_minimal() +
  scale_x_discrete(labels = c("Control", "Treatment"))

interviewed_rate_4b

# Employed Now Rate

employed_rate_4c <- ggplot(df_stats, aes(x = factor(condition), y = employed_now_rate)) +
  geom_bar(stat = "identity", fill = "lightgreen", color = "black", width = 0.7) +
  geom_errorbar(aes(ymin = employed_now_rate_lo, ymax = employed_now_rate_hi), width = 0.2, color = "black") +
  geom_text(aes(label = sprintf("%.2f%%", employed_now_rate), y = employed_now_rate_hi + 2), vjust = 0) +
  annotate("text", x = 1.5, y = max(df_stats$employed_now_rate_hi) + 5, label = paste("p-value =", round(p_value_employed_now, 3)), size = 4) +
  labs(y = "% Employed Now", x = "") +
  theme_minimal() +
  scale_x_discrete(labels = c("Control", "Treatment"))

employed_rate_4c

# Driving Rate
driving_rate_4d <- ggplot(df_stats, aes(x = factor(condition), y = driving_rate)) +
  geom_bar(stat = "identity", fill = "lightpink", color = "black", width = 0.7) +
  geom_errorbar(aes(ymin = driving_rate_lo, ymax = driving_rate_hi), width = 0.2, color = "black") +
  geom_text(aes(label = sprintf("%.2f%%", driving_rate), y = driving_rate_hi + 2), vjust = 0) +
  annotate("text", x = 1.5, y = max(df_stats$driving_rate_hi) + 5, label = paste("p-value =", round(p_value_driving, 3)), size = 4) +
  labs(y = "% Driving", x = "Condition") +
  theme_minimal() +
  scale_x_discrete(labels = c("Control", "Treatment"))

driving_rate_4d
```

\newpage
@fig-national_survey_misperception is made to illustrate the result of the first national survey. There are two vertical lines that represent the actual proportion of respondents who agreed with the statement in @fig-national_survey_misperception, showing more than 80% of men underestimated the true proportion. The black line is for the control group and the red dashed line is for the treatment group. The lines reflect perceptions of participants about others' beliefs. The graph shows the support level of treatment group is mostly higher than control. This figure underscores a significant discrepancy between personal beliefs and perceived societal norms. The larger sample size of this national survey strengthens the evidence that men privately support women's employment outside the home more than they assume others do, suggesting widespread misperceptions about societal norms.

```{r}
#| message: false
#| echo: false
outside_value <- na.omit(online_survey_clean$c_outside_guess_frac)
group <- rep("Control", length(outside_value))
c_outside <- cbind(outside_value, group)
c_outside <- data.frame(c_outside)

outside_value <- na.omit(online_survey_clean$t_outside_guess_frac)
group <- rep("Treatment", length(outside_value))
t_outside <- cbind(outside_value, group)
t_outside <- data.frame(t_outside)
data_frame <- rbind(c_outside, t_outside)

c_mean <- mean(online_survey_clean$c_outside_mean)/100
t_mean <- mean(online_survey_clean$t_outside_mean)/100

data_frame <- data.frame(data_frame)
data_frame$group <- as.factor(data_frame$group)
data_frame$outside_value <- as.numeric(data_frame$outside_value)
```

```{r}
#| message: false
#| echo: false
#| label: fig-national_survey_misperception
#| fig-cap: Misperceptions about Others’ Beliefs (National Survey)
data_frame |>
  ggplot(aes(x = outside_value)) +
  theme_minimal() +
  stat_ecdf(aes(color = group, linetype = group), linewidth = 0.5) +
  scale_y_continuous(name = "Cumulative Probability", breaks = c(0,.2,.4,.6,.8,1), limits = c(0,1)) +
  scale_x_continuous(name = "Share of other agreeing", breaks = c(0,.2,.4,.6,.8,1), limits = c(0,1)) +
  theme(legend.position = "bottom", legend.title = element_blank(),
        legend.direction = "vertical",
        aspect.ratio = 0.5,
        legend.box.background = element_rect(color = "black", linewidth = 0.5)) +
  scale_linetype_discrete(labels = c("Control (perceptions about others' answers)", "Treatment (Perception about others' beliefs)")) +
  scale_colour_manual(values = c("black","tomato4"), labels = c("Control (perceptions about others' answers)", "Treatment (Perception about others' beliefs)")) +
  geom_vline(aes(xintercept = c_mean), col = "black")+
  geom_vline(aes(xintercept = t_mean), col="tomato4", linetype ="longdash") +
  annotate(geom = "text",
           label = c("True proportion\n(treatment)","True proportion\n(control)"),
           x = c(0.8, 1),
           y = 0.15,
           vjust = -1,
           hjust = 1,
           size = 2,
           col = c("black","black"))
```

# Discussion
## Impact of Informational Intervention: 
The research findings have highlighted the importance of correcting misperceived social norms, not only to increase women's participation in the labor market but also to make progress in broader social norms and attitudes towards women's rights and autonomy. The study shows that interventions aimed at challenging and changing societal norms can have a significant impact on women's economic participation and social development. For instance, the study found that after an information intervention, more husbands were willing to sign up their wives for driving lessons, indicating a shift towards more progressive attitudes regarding women's mobility and independence. These changes can contribute to economic development by increasing the labor force and promoting gender equality. The researchers suggest that similar interventions could be effective in other contexts where misperceptions about social norms limit women's economic participation. This approach can offer valuable insights for policymakers and organizations aiming to promote gender equality and economic development, emphasizing the role of accurate information in challenging and changing societal norms. In today's world, digital platforms and social media hold a significant impact on the formation of public opinion and social norms. By making strategic use of these technologies, we can effectively disseminate accurate information and correct any misconceptions. We can leverage online platforms to share data on societal attitudes towards WWOH. This will enable interventions to reach a much wider audience, facilitating a more rapid and widespread change in perceptions.

## Broader Social Proof Implications
It is possible to overcome harmful norms in cultural and societal settings by using social proof.

One study @Counternormative found that communicating the increasing popularity of environmentally friendly behaviors encouraged individuals to adopt these behaviors themselves, demonstrating the power of social proof in effecting change in environmental contexts. This suggests that highlighting positive changes in societal behaviors can be effective in encouraging the adoption of new, more beneficial norms and practices, rather than focusing on static norms.

Another study @OrganizedDiffusio has evaluated the impact of "organized diffusion" on community-based health and women's empowerment programs. This approach involves encouraging individuals who have received training or information to share it with others, thus spreading the benefits more widely. The study, which looked at programs in Mali, Nepal, and Nigeria, found that implementing this method can lead to significant changes in social norms without incurring additional costs.

These studies show that by understanding and addressing specific normative beliefs underlying societal issues, policymakers and practitioners can design targeted interventions that utilize social proof effectively. By presenting evidence of changing norms and behaviors within a community, we can foster a more inclusive, equitable, and sustainable society.

Integrating these studies enhances our understanding of how social proof can be applied across different issues and cultural contexts to promote positive social change.

## Inverse Pattern of Saudi female labor force participation rate during pandemic
According to the Gender Data Portal from The World Bank, the global average male labor force participation had a decrease of 1.6%, reaching 58.6% in 2020, followed by a rebound to 59.5% in 2021. In fact, this trend was observed across various countries, irrespective of gender.

For example, the Canadian female labor force participation rates in 2019, 2020 and 2021 were 61.4%, 59.5%, 60.6%, respectively, and the Canadian male labor force participation rates for the same years were 70.1%, 68.6%, 69.6%. A similar pattern is discernible in India, where the female labor force participation rates in 2019, 2020 and 2021 were 23.7%, 22.7%, 23%, while corresponding male rates stood at 73.6%, 72.3%, 72.7% [@worldbankLaborForce].

Interestingly, Saudi Arabia exhibited an inverse pattern. While the Saudi male labor force participation rate remained relatively stable, the female rate, along with that of the United Arab Emirates, increased in 2020 and then decreased in 2021. To be more specific, the Saudi female labor force participation rate had a notable increase of 4.9% (nearly one-fifth of the total), reaching 29.5% in 2020, and then declining to 27.8% in 2021. Similarly, female rate in United Arab Emirates increased by 4.2% to 61% in 2020, but decreased to 52.8% in 2021 [@worldbankLaborForce].

The observed fluctuation in world labor force participation rates, particularly Saudi Arabia, leads to speculation on the underlying factors driving these trends. The impact of quarantine is a significant factor, as individuals working from home remained mostly unaffected, unlike those in traditional office settings. This is likely the reason behind the abnormal employment pattern among Saudi women at the start of the pandemic. Moreover, the contrasting pattern in Saudi female labor force participation rate, compared to the world trend, highlights the phenomenon  that many employed females tend to work at home, either willingly or due to traditional social norms. The 2020 surge could indicate potential willingness among Saudi women to explore employment opportunities.

## Bias, Ethics and Weaknesses
Throughout the survey and experiment, all participants granted permissions for the collection of data. After gathering information, the experimenters ensured that all responses remain confidential and anonymous, so the data we used for replication in this paper was also anonymous, with only the last three digits of their phone numbers retained. As a result, there are no ethical concerns during the process and in the paper.

The authors, Leonardo Bursztyn, Alessandra L. González, and David Yanagizawa-Drott, implemented many modifications to the questions and weights in an attempt to minimize bias in the survey. However, it is acknowledged that bias is always a challenge when conducting experiments. The main experiment survey has a participant pool of 500 married men from Riyadh [@MisperceivedSocialNorm]. Since Riyadh is the capital and the largest city in Saudi Arabia, the sample size might be relatively small and potentially biased. Considering the economy in Saudi Arabia and Riyadh, the reward offered, typically a five-dollar or twenty-dollar gift card, could be another factor contributing to the weakness in both the original paper and our paper.

The paper focuses on married men aged 18 to 35 with some level of post-secondary education. It is noted that the enrolment rate among students aged 20-24 in secondary program is relatively high in Saudi Arabia [@EducationGPS], so setting the age range from 18 to 35 might be a weakness in the experiment. There is a potential correlation between individuals who did not participate in the follow-up and those who do not support Women Working Outside the Home, also introducing the possibility of bias.

## Next steps
Although there is a customary practice of husbands' guardianship with decision-making authority in Saudi Arabia, gaining insights into the opinions of Saudi women regarding employment and working outside the home is essential for future study. The authors mentioned that ‘expanding the scale and observing how information spreads in networks, and how it affects a large set of outcomes is an important topic for future work’ [@MisperceivedSocialNorm]. Considering the potential biases and the experiment result claiming that correcting misperceived beliefs increase husbands’ support in WWOH, future research could greatly benefit from the extension of scale to enhance generalizability and actual effects on female employment in Saudi society.

Figure 9 in the original paper reveals that males who often engage in discussions about the WWOH have perceived proportion of support only four percent smaller than the actual proportion. For those who rarely or very rarely discuss about the WWOH, their perceived proportion of support for WWOH is around forty percent smaller than the actual proportion. If the lack of communication facilitates misperceptions, one would expect a correlation between the frequency of communication and perceptions [@MisperceivedSocialNorm]. Consequently, future studies could delve into strategies in preventing misperceived opinions from influencing social norms negatively, and explore decision-making in the presence of collective misperceptions.

\newpage

# References